Reproducible Research
=====================

## Loading and Prepocesiing data

```{r, echo=TRUE, results='hide'}

activity<-read.csv("activity.csv")
noNA<-na.omit(activity)
```

# What is the total number of steps taken per day?

```{r, echo=TRUE}
stepsPerDay<- aggregate(steps~date, noNA, sum)
hist(stepsPerDay$steps, breaks=10, xlab="Steps per Day")
rmean<-mean(stepsPerDay$steps)
rmedian<-median(stepsPerDay$steps)
```

The mean is `r rmean` and median is `r rmedian`

## What is the average daily activity pattern?

```{r, echo=TRUE}
stepsByInterval<- aggregate(steps~interval, noNA, mean)
plot(stepsByInterval$interval, stepsByInterval$steps, type="l", xlab="5 minute Interval", ylab="Average number of steps")
maxInterval<- stepsByInterval[which.max(stepsByInterval$steps),1]
```

The 5-minute interval with the maximum average steps is `r maxInterval`

## Imputing missing values

```{r, echo=TRUE}
missingValues<- sum(!complete.cases(activity))
```

The number or missing values in the dataset is `r missingValues`

Now, filling in missing values with the mean 5-minute interval values
```{r, echo=TRUE}
newData<- activity
for(i in 1:nrow(newData)){
if(is.na(newData$steps[i])){
  newData$steps[i]<-stepsByInterval[which(stepsByInterval$interval==newData$interval[i]),]$steps
}}
```

Recounting total steps per day and creating histogram

```{r, echo=TRUE}
newStepsPerDay<- aggregate(steps~date, newData, sum)
hist(newStepsPerDay$steps, breaks=10, xlab="Adjusted Steps per Day")
newMean<- mean(newStepsPerDay$steps)
newMedian<-median(newStepsPerDay$steps)
meanDiff<- rmean-newMean
medianDiff<- rmedian-newMedian
```

the imputed data mean is `r newMean`
the imputed data median is `r newMedian`

the difference in mean values is `r meanDiff`
the difference in median values is `r medianDiff`

## Are there differences in activity patterns between weekdays and weekends?

```{r,echo=TRUE}
newData$date<- as.Date(newData$date)
newData$dow<- as.factor(ifelse(weekdays(newData$date)%in% c("Saturday","Sunday"), "weekend", "weekday"))
library(lattice)
xyplot(newData$steps~newData$interval|newData$dow, layout=c(1,2), type="l", xlab="Interval", ylab="Steps")
```

